---
title: "Exploratory Data Analysis with sparklyr and R"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


## Visualizing Scatterplots

We can take our Spark DataFrame and use it for visualizations. There are some libraries that can directly plot Spark DataFrames. However, visualizing a huge collection of points is rarely efficient or effective. It's better to summarise first and then visualize. Here we will use Spark SQL to summarise our data, and then we'll collect it, and then visualize it.

We'll use the popular `ggplot2` package for visualization.

```{r delays}

delay <- airWeatherDF %>% 
  group_by(tailNum) %>%
  summarise(count = n(), 
            dist = mean(distance), 
            delay = mean(as.numeric(arrDelayNew))) %>%
  filter(count > 20, dist < 2000, !is.na(delay)) %>%
  collect

# plot delays
library(ggplot2)
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area(max_size = 2)

```


## Visualizing Delays as Bar Charts

Let's see if we can visualize delays using bar charts by carrier and day of week.

We will again summarise using `sparklyr`, this time aggregating over the `dayOfWeek` column and the `uniqueCarrier` column. We'll convert the `dayOfWeek` integer into a factor variable.

```{r barcharts-delays}


ave_delay_carrier <- airWeatherDF %>% group_by(dayOfWeek, uniqueCarrier) %>% 
  summarise(aveDelay = mean(as.numeric(arrDelayNew))) %>% 
  collect


library(forcats)

ave_delay_carrier <- ave_delay_carrier %>% ungroup() %>% 
  mutate(dayOfWeek = fct_recode(factor(dayOfWeek), 
                                "Monday" = "1",
                                "Tuesday" = "2",
                                "Wednesday" = "3",
                                "Thursday" = "4",
                                "Friday" = "5",
                                "Saturday" = "6",
                                "Sunday" = "7"))

ggplot(ave_delay_carrier) + 
  geom_bar(aes(x = dayOfWeek, 
               y = aveDelay,
               fill = uniqueCarrier),
           stat = 'identity') +
  facet_wrap(~uniqueCarrier) + 
  coord_flip() + theme_minimal() + guides(fill = FALSE) +
  labs(title = "Average Delay by Carrier and Day of Week",
       subtitle = "(minutes)",
       xlab = "Minutes",
       ylab = "Carrier")



```

## Fancy Interactive Visualizations Using WebGL

R has an DSL for interacting with `html`. We can create javascript based visualizations directly with R, and they'll render natively in our IDE session and output. Let's join some airport geospatial data with our Spark DataFrame, and then visualize the route traveled.

```{r airport-geo}


# WebGL Routes Visualization ----------------------------------------------

# system("wget https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat")

airports <- read.table("airports.dat", header = F, sep = ",")

airportsDF <- copy_to(sc, airports, "airports")
airportsDF <- rename(airportsDF,
                     airportID = V1,
                     airportName = V2,
                     City = V3,
                     Country = V4, 
                     airportIATA = V5, 
                     airportICAO = V6, 
                     latitude = V7,
                     longitude = V8,
                     altitude = V9,
                     timezone = V10,
                     DST = V11,
                     TZdb = V12,
                     Type = V13,
                     Source = V14)

origins <- left_join(airWeatherDF %>% select(originAirportId),
                     airportsDF %>% select(airportID, airportName, longitude, latitude),
                     by = c("originAirportId" = "airportID"))

dest <- left_join(airWeatherDF %>% select(destAirportId), 
                  airportsDF %>% select(airportID, airportName, longitude, latitude),
                  by = c("destAirportId" = "airportID"))

origin_tbl <- origins  %>% collect
dest_tbl <- dest  %>% collect
codes <- bind_cols(origin_tbl, dest_tbl) %>% repair_names
codes <- codes %>% filter(!is.na(airportID), !is.na(airportID1))

flights <- codes %>% select(latitude, longitude, latitude1, longitude1)

```

Now that we have the coordinates of the origin and destination, we can visualize it using a global map:

```{r webgl-globe}

library(threejs)

earth <- system.file("images/world.jpg",  package="threejs")

globejs(img=earth, arcs=flights,
        arcsHeight=0.3, arcsLwd=2, arcsColor="#ffff00", arcsOpacity=0.15,
        atmosphere=TRUE)
```


## Disconnect from Spark Session

When you're done with your Spark Session, make sure to disconnect:

```{r disconnect}

spark_disconnect(sc)

```

